Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-03-03 writing/transect_paper.Rmd
Add class taxonomy chart, switch back to Molecular Ecology

---
title: "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
author: 
  - Brendan Furneaux^1^,
    Mohammad Bahram^2,3^,
    Anna Rosling^4^,
    Nourou S. Yorou^5^,
    Martin Ryberg^1^
  - ^1^Program in Systematic Biology,
    Department of Organismal Biology,
    Uppsala University,
    Uppsala, Sweden
  - ^2^Department of Ecology,
    Swedish University of Agricultural Sciences,
    Uppsala, Sweden
  - ^3^Institute of Ecology and Earth Sciences,
    University of Tartu,
    Tartu, Estonia
  - ^4^Program in Evolutionary Biology,
    Department of Ecology and Genetics,
    Uppsala University,
    Uppsala, Sweden
  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
    LEB,
    University of Parakou,
    Parakou, Benin
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography:
  - "all.bib"
  - "R.bib"
csl: new-phytologist.csl
always_allow_html: true
output:
  bookdown::pdf_document2:
    fig_crop: yes
    keep_tex: yes
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
  bookdown::word_document2: 
    keep_md: yes
---

@@

# Methods

@@

(ref:taxa) Color and size of nodes represent the fractional read abundance of all ASVs belonging to that taxonomic group.
Color and thickness of branches represent the fractional ASV richness of that taxonomic group.
Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
ASV richness and read abundance are displayed as the mean across sequencing runs.
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.

@@

@@

# Conclusion

@@

# Data Availability Statement

# Author contributions

# Acknowledgements

@@

# Bibliography {-}

@@

@@

(ref:heattree-length-compare) Red nodes (edges) represent increased read (ASV) count for the short amplicon library, while blue-green nodes (edges) represent increased read (ASV) count for the long amplicon library.
Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.

@@

(ref:ecm-heattree) Red edges/nodes represent increased read/ASV count for the short amplicon library, while teal edges/nodes represent increased read/ASV count for the long amplicon library.
Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.

>>>>

---
title: "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
author: 
  - Brendan Furneaux^1^,
  -  Mohammad Bahram^2,3^,
  - Anna Rosling^4^,
  - Nourou S. Yorou^5^,
  - Martin Ryberg^1^
  - ^1^Program in Systematic Biology,
    Department of Organismal Biology,
    Uppsala University,
  - Uppsala, Sweden
  - ^2^Department of Ecology,
    Swedish University of Agricultural Sciences,
    Uppsala, Sweden
  - ^3^Institute of Ecology and Earth Sciences,
    University of Tartu,
    Tartu, Estonia
  - ^4^Program in Evolutionary Biology,
    Department of Ecology and Genetics,
    Uppsala University,
  - Uppsala, Sweden
  - ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
    LEB,
  - University of Parakou,
    Parakou, Benin
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography:
  - "all.bib"
  - "R.bib"
csl: molecular-ecology.csl
always_allow_html: true
output:
  bookdown::pdf_document2:
    fig_crop: yes
    keep_tex: yes
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
  bookdown::word_document2: 
    keep_md: yes
---

@@

# Materials and Methods

@@

(ref:taxa) Color and size of nodes represent the fractional ASV richness of that taxonomic group.
Color and thickness of branches represent the fractional read abundance of all ASVs belonging to that taxonomic group.
Diverse but relatively rare groups are thus shown as large, dark-colored nodes on relatively thin, light-colored branches (e.g., Dothideomycetes), while common but relatively non-diverse groups are shown as small, light-colored nodes on relatively thick, dark-colored branches (e.g., Russulales).
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
ASV richness and read abundance are displayed as the mean across sequencing runs.
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.

@@

(ref:fungi-comm-s) Taxonomic composition of fungal community at the class level

(ref:fungi-comm) Values represent the fraction of all ASVs, OTUs, or reads which were assigned to kingdom Fungi.
Assignments based on PHYLOTAX.
Classes which represented less than 2% of reads, OTUs, and ASVs in all datasets are grouped together as "other".

```{r fungi-comm, fig.cap = "(ref:fungi-comm-s). (ref:fungi-comm)", fig.scap = "(ref:fungi-comm-s)"}
rdd(taxon_reads) %>%
  filter(
    kingdom == "Fungi",
    !(Algorithm == "Consensus" & region == "ITS"),
    Algorithm == "PHYLOTAX"
    ) %>%
  group_by(tech, amplicon) %>%
  mutate(reads = reads/sum(reads), ASVs = 1/n()) %>%
  pivot_longer(cols = c("reads", "ASVs"), names_to = "type", values_to = "reads") %>%
  bind_rows(
    filter(., type == "ASVs", label == OTU) %>%
      mutate(label = OTU) %>%
      group_by(tech, amplicon) %>%
      mutate(reads = 1/n(), type = "OTUs")
  ) %>%
  taxon_plot(
    rank = class,
    datasets = datasets,
    cutoff = 0.02,
    y = reads,
    x = type,
    facets = vars(tech, amplicon)
  ) +
  ylab("Fraction of Community") +
  xlab(NULL)
```

@@

## Conclusion

@@

# Acknowledgements {-}

@@

# References {-}

@@

# Data Availability Statement {-}

# Author contributions {-}

@@

(ref:heattree-length-compare) Color of nodes indicates variation in read abundance, while color of edges represents variation in ASV richness.
In both cases red represents increased prevalence in the short amplicon library, while blue-green represents increased prevalence in the long amplicon library.
Values are log~10~ of the ratio of mean values across sequencing runs for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Only nodes with a log read abundance ratio greater than 0.5 (abundance ratio > 3.2) are labeled.
Refer to Figure\ \@ref(fig:taxa) for complete taxon labeling.

@@

(ref:ecm-heattree) Red edges (nodes) represent increased read (ASV) count for the short amplicon library, while blue-green edges (nodes) represent increased read (ASV) count for the long amplicon library.
Values are log~10~ of the ratio of mean values across all sequencing technologies for each amplicon.
Groups which are not represented by at least 1% of reads or 1% of ASVs in any dataset are collapsed into nodes labeled "*".
ASVs which could not be further assigned are labelled "?".
Taxonomic assignment is by strict consensus of multiple ITS databases (Unite, Warcup) and assignment algorithms (RDPC, SINTAX, IDTAXA).
The RDP LSU database was not included, to ensure consistent assignment between long and short amplicon libraries.
